---
title: "ESM 206 Lab 5"
author: "Ruiheng Jin"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## Objectives:

- Getting counts for different groups
- Use {lubridate} package -- elucidate dates
- Find confidence intervals & do t-tests with t.test()
- Heatmap with geom_tile()

```{r Attach packages, include=FALSE}
# attach packages
library(tidyverse)
library(janitor)
library(here)
```

```{r Import data and initial cleaning}
# import lobster data

lobster_abun <- read_csv(here::here("data", "lobster_abundance.csv"),
                         na = "-99999") %>% 
  clean_names()

# date column is stored as characters
##### lobster_count - not tidy, we want each lobster have its own row #####
# NA are stored as -99999
```

Use the dplyr::uncount() function to expand a frequency variable

```{r Use the dplyr::uncount() function to expand a frequency variable}
lobster_tidy <- lobster_abun %>%
  tidyr::uncount(lobster_count)
```
Now each lobster has its own row, it's in tidy format.


Exploratory data visualization
```{r Exploratory data viz}
ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site), 
              width = .2,
              alpha = .3)

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(fill = site, color = site))

# histogram
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site), show.legend = FALSE) +
  facet_wrap(~site, scales = "free")
##### Have diff scales for each facet #####


# QQ plot 
ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = .5) + 
  facet_wrap(~site, scales = "free")
```

Use "lubridate" function to parse dates and times, add new column with mutate() that contains date as actual dates
```{r lubridate function}
##### need to know the function of the date, MDY? YMD? choose function accordingly #####
lobster_date <- lobster_tidy %>% 
  mutate(date_new = lubridate::mdy(date))
# check the class of new col, it's "date"

# what if we want to group by month?

```

What if we want to group by month?
Parse "date" column to get different pieces separated
```{r Parse "date" column to get different pieces separated}
##### lubridate::month, label = TRUE #####
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_months = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
```

Count lobsters by diff groupings
COUNT NUMBER OF LOBSTERS BY YEAR AND MONTH

```{r Counting things}
lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_months)

lobster_ym
# it counted the number of observations by these groupings

# try a diff grouping
lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)
lobster_ysite

lobster_site <- lobster_parse_date %>% 
  count(site)
lobster_site


# If we want other summary statistics in addition to count
##### function n() #####
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
  )
```

Confidence intervals and t-tests.

Use the 't.test()' function to find CIs and perform t tests

```{r t test and confidence intervals}

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)
# default is one sample t test, see if it's significantly different from 0

```

Two-sample t-test
Is there a sig diff in lobster length between Naples and Mohawk?
```{r two sample t-test}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)

mohk_napl_ttest


```
Use ?t.test() to check values stored in the test, to reference them

Here's how to reference test statistical results
Mean lobster size differened sig b/n Mohawk and Naples reefs:
(t('r mohk_napl_ttest$parameter') = 'r mohk_napl_ttest$statistic')

maybe an easier way...
```{r another way to do 2 sample ttest}
lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)
ttest_2
```

Make a geom_tile heatmap
```{r}

ggplot(lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```




















